### 2016-03-22 reply_to_reviewers.md

Dear Professor Freckleton,

We thank the associate editor and reviewer for their comments, which have
improved our manuscript and the software it describes. Below we provide detailed
replies to specific comments.

The developers working on the Open Tree of Life APIs have informed us that a new
version of the APIs is going to be released in the next few weeks. We don't
expect these changes to impact what we describe in the manuscript, but small
tweaks to the package might be needed.

We look forward to hearing back from you.

  François Michonneau on behalf of Joseph Brown and David Winter

@@

We have expanded this section a little, hopefully making the context clearer:

_Before a researcher can use the Open Tree to retrieve relationships among a set
of taxa, they first need to match the taxon names in their dataset with records
in the Open Tree Taxonomy (OTT). The TNRS allows users to search for taxon
names, and retreive identifiers for each matching taxon. We demonstrate the use
of the TNRS within `rotl` by searching for taxonomic records associated
with several model organisms._

>l. 118: "henihomonyms" may warrant a definition or replacement with
>something simpler.

replaced with "_names that may represent valid taxa in more than one code._"

@@

Not at present, but it's currently in progress. See:
https://github.com/ropensci/taxize/issues/517

## Reviewer 1

> One thing I thought was missing was a discussion about the importance of rolt
> to the field of ecology.

We added a paragraph (l. 60-70) that describes how `rotl` can be used as a
data source for community phylogenetics. We also mention community ecology in
the concluding remarks (l. 202-210).

@@

> 1) Can user generated trees be easily incorporated into an output tree from
> rotl? Perhaps by implementing some of the algorithms used to integrate the
> studies that comprise the OTL? For example, if I generated a tree for a genus
> that was not yet in the OTL, could I add it into its family in the rotl
> interface, or would I have to go through the curator interface?

It's an interesting idea. If a researcher wants to include a better resolved
tree than what is available in Open Tree, it is relatively straightforward to
swap a monophyletic group by another one using the functions provided by
`ape`. However, we feel that providing this kind of functionality falls beyond the
scope of our package, whose main purpose is to provide an interface to the OTL
API.

> 2) I was a bit confused about why the authors chose to include the GBIF
> example in the appendix, but not the other vignettes. Why not just add it as
> another vignette, or include the other vignettes as appendices as well? A
> small detail perhaps, but this makes the paragraph before the concluding
> remarks a bit vague.

We included the code for the GBIF example as an appendix as it seemed too long
to be included in the text of the manuscript, and not long enough to warrant its
own vignette. We didn't include the other vignettes as appendices to ensure that
we can update the code they include when the API changes.

@@

> 3) How might rolt be combined with trait databases? This was briefly mentioned
> at the end of the vignette “Connecting data to Open Tree trees”, but might be
> mentioned in the manuscript as well (for example, in the subsection “How does
> rotl fit into the R package ecosystem?”).

We added a paragraph to highlight that `rotl` can be used to provide a
phylogenetic context to available trait data (l. 189-191).

@@

> Page 2, paragraph 3, lines 39-41: This sentence is a bit awkward—perhaps
> remove the comma?

Done

> Page 3, paragraph 1, line 58: Other recently developed R packages that relate
> to rotl might be mentioned, including taxize (taxonomy congruence),
> phyloGenerator (generate phylogenies for ecologists) and pez (for calculating
> suites of metrics for community phylogenetics).

We added a paragraph about the use of phylogenetics in community ecology,
including a description of R packages that can be used for these analysis
(l. 65-70).

@@

> Page 4, point #2: So not all studies that can be queried through the API were
> included in the OTL? Perhaps this can be explained a bit more here (or in the
> “Getting trees from studies” subsection).

We added the number of studies that have been submitted through the curator app,
and the number of studies that contribute to the synthetic tree to clarify this
point: "_As of March 2016, the Open Tree of Life project stores 7755 trees from
3399 studies (each having between 0 and 61 trees), and 477 of these trees are
used to assemble the synthetic tree._"

@@

> Page 9, line 150: What type of metadata can be extracted from a study? Is it
> just the citation (which is the example given in “How to use rotl?” vignette)?

There is a lot more information associated with each study. We re-worded the
paragraph to mention some of them: "_Metadata about the study (e.g., citation
information, information about the curator for the study, and other technical
information regarding the import of this study) can be obtained using the
function `get_study_meta`_."

@@

> Page 10, paragraph 2, line 169: “other data” is a bit vague. Integration with
> other data, for example species distribution data, is demonstrated in the
> appendix (but not the main manuscript)

We re-wrote this paragraph to be more specific about the type of data that can
be linked to the phylogenies storred in `rotl` (l. 168-169).

### 2016-03-22 manuscript.Rnw
fix line breaks

In ecology, the development of the field of community phylogenetics
\citep{Webb2002} has created a need for researchers to have access to the
evolutionary relationships of species making up communities. The relative
contributions of the role of the environment (e.g., habitat filtering), and of
biotic interactions (e.g., competitive exclusion), are inferred from the
distribution of taxa on a phylogeny composed from species occurring at larger
spatial scale. R packages and other software have been developed to generate
phylogenies from species lists using taxonomic information or DNA sequences
\citep[e.g.,][]{Webb2008,Pearse2013}. These phylogenies can then be used for
community phylogenetics analyses \citep[e.g.,][]{Pearse2015},
\citep{Kembel2010}, but they are often incomplete or not resolved enough. As OTL
becomes more comprehensive, and its taxonomic resolution increases, it could
become a valuable resource for ecologists seeking to use phylogenetic
information in their research.

>>>>

In ecology, the development of the field of community phylogenetics
\citep{Webb2002} has created a need for researchers to have access to the
evolutionary relationships of species making up communities. The relative
contributions of the role of the environment (e.g., habitat filtering), and of
biotic interactions (e.g., competitive exclusion), are inferred from the
distribution of taxa on a phylogeny composed from species occurring at larger
spatial scale. R packages and other software have been developed to generate
phylogenies from species lists using taxonomic information or DNA sequences
\citep[e.g.,][]{Webb2008,Pearse2013}. These phylogenies can then be used for
community phylogenetics analyses \citep[e.g.,][]{Pearse2015, Kembel2010}, but
they are often incomplete or not resolved enough. As OTL becomes more
comprehensive, and its taxonomic resolution increases, it could become a
valuable resource for ecologists seeking to use phylogenetic information in
their research.
